---
title: "ADA Combined Code"
author: "Amanda Gilbert & Pranav Nandan & Victor Martinez-Cassmeyer"
date: "4/6/2020"
output: html_document
---




# Load Data
```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)
library(dplyr)
library(readr) #for read txt file
library(survival) #for calculating KM values
library(survminer)#for pairwise diffs
library(ggfortify) #for KM curves
library(readxl) # for reading in excel file
library(ggplot2) # for plotting KM curve
library(haven) #for read_csv

covid_org5 <- read_csv("https://raw.githubusercontent.com/beoutbreakprepared/nCoV2019/master/dataset_archive/covid-19.data.2020-04-14T075627.csv")

covid.clean <- covid_org5 %>%
  mutate(outcomec = recode(covid_org5$outcome,
                         "death" = "died",
                         "died" = "died",
                         "dead" = "died",
                         "discharge" = "discharge",
                         "discharged" = "discharge",
                         "Discharged" = "discharge",
                         "Discharged from hospital" = "discharge",
                         "recovered" = "discharge",
                         "not hospitalized"= "not hospitalized",
                         "released from quaratine" = "released from quarantine",
                         "severe" = "severe",
                         "stable" = "stable",
                         "alive" = "alive",
                         "Alive" = "alive",
                         "Receiving Treatment" = "hospitalized",
                         "Critical condition" = "hospitalized",
                         "critical condition, intubated as of 14.02.2020" = "hospitalized",
                         "Dead" = "died",
                         "Death" = "died",
                         "Died" = "died",
                         "Deceased" = "died",
                         "Recovered" = "discharge",
                         "recovering at home 03.03.2020" = "discharge",
                         "severe illness" = "severe",
                         "stable condition" = "stable",
                         "Stable" = "stable",
                         "Symptoms only improved with cough. Currently hospitalized for follow-up." = "hospitalized",
                         "treated in an intensive care unit (14.02.2020)" = "hospitalized",
                         "Under treatment" = "hospitalized",
                         "unstable" = "hospitalized")) %>%
  mutate_all(as.character) %>% 
  #combine date_onset_symptoms and "date_admission_hospital" into onset_date
  mutate(onset_date = ifelse(is.na(date_onset_symptoms),date_admission_hospital , date_onset_symptoms)) %>%
  #combined onset_date with date_confirmation to become onsetdate
  mutate(onsetdate = ifelse(is.na(onset_date),date_confirmation , onset_date)) %>%
  #select wanted variables  
  select(ID, age, sex, country, onsetdate, outcomec, date_death_or_discharge) %>%
  mutate(country_binary = ifelse(country == "China", "china", "country_other")) %>%
  # make binary event variable
  mutate(event = ifelse(outcomec == "died", 1, 0)) %>%
  # make time variable
  # 1. random row was a range, corrected it to earliest point
  mutate(onsetdate = ifelse(onsetdate == "25.02.2020 - 26.02.2020", "25.02.2020", onsetdate)) %>%
  # 2. correct date formats
  mutate(onsetdate_correct = as.Date(gsub('[.]', '-', onsetdate), "%d-%m-%Y")) %>%
  mutate(datedeathordischarge_correct = as.Date(gsub('[.]', '-', date_death_or_discharge), "%d-%m-%y")) %>%
  # 3. subtract dates and record output in days
  mutate(time = difftime(datedeathordischarge_correct, onsetdate_correct, units = "days"), time = as.numeric(time)) %>%
  # filter times that are negative (due to testing delays)
  filter(time >= 0) %>%
  mutate(age_cat = recode(age,
                         "1" = "35 or lower",
                         "4" = "35 or lower",
                         "7" = "35 or lower",
                         "9" = "35 or lower",
                         "8" = "35 or lower",
                         "13" = "35 or lower",
                         "16" = "35 or lower",
                         "20"= "35 or lower",
                         "21" = "35 or lower",
                         "22" = "35 or lower",
                         "23" = "35 or lower",
                         "24" = "35 or lower",
                         "25" =  "35 or lower",
                         "26" = "35 or lower",
                         "27" = "35 or lower",
                         "28" = "35 or lower",
                         "29" = "35 or lower",
                         "20-29" = "35 or lower",
                         "30" = "35 or lower",
                         "31" = "35 or lower",
                         "32" = "35 or lower",
                         "33" = "35 or lower",
                         "34" = "35 or lower",
                         "35" = "35 or lower",
                         "36" = "36-65",
                         "37" = "36-65",
                         "38" = "36-65",
                         "39" = "36-65",
                         "40" = "36-65",
                         "41" = "36-65",
                         "42" = "36-65",
                         "43" = "36-65",
                         "44" = "36-65",
                         "45" = "36-65",
                         "46" = "36-65",
                         "47" = "36-65",
                         "48" = "36-65",
                         "49" = "36-65",
                         "50" = "36-65",
                         "51" = "36-65",
                         "52" = "36-65",
                         "53" = "36-65",
                         "54" = "36-65",
                         "55" = "36-65",
                         "56" = "36-65",
                         "57" = "36-65",
                         "58" = "36-65",
                         "59" = "36-65",
                         "50-59" = "36-65",
                         "60" = "36-65",
                         "61" = "36-65",
                         "62" = "36-65",
                         "63" = "36-65",
                         "64" = "36-65",
                         "65" = "36-65",
                         "66" = "66 and older",
                         "67" = "66 and older",
                         "68" = "66 and older",
                         "69" = "66 and older",
                         "70" = "66 and older",
                         "71" = "66 and older",
                         "72" = "66 and older",
                         "73" = "66 and older",
                         "74" = "66 and older",
                         "75" = "66 and older",
                         "76" = "66 and older",
                         "77" = "66 and older",
                         "78" = "66 and older",
                         "79" = "66 and older",
                         "70-79" = "66 and older",
                         "80-" = "66 and older",
                         "81" = "66 and older",
                         "82" = "66 and older",
                         "83" = "66 and older",
                         "94" = "66 and older",
                         "84" = "66 and older",
                         "85" = "66 and older",
                         "86" = "66 and older",
                         "87" = "66 and older",
                         "88" = "66 and older",
                         "89" = "66 and older",
                         "80" = "66 and older",
                         "80-89" = "66 and older",
                         "92" = "66 and older",
                         "90-99" = "66 and older",
                         "95" = "66 and older",
                         "0.25" = NA_character_,
                         "0.5" = NA_character_,
                         "60-69" = NA_character_)) %>%
  mutate(country_binary2 = ifelse(country_binary == "china", 1, 0)) %>%
  mutate(age_cat2 = ifelse(age_cat == "35 or lower", 0, 1)) %>%
  mutate(age_cat2 = ifelse(age_cat == "36-65", 1, age_cat2)) %>%
  mutate(age_cat2 = ifelse(age_cat == "66 and older", 2, age_cat2)) %>%
  mutate(sex2 = ifelse(sex == "male", 1, 0)) %>%
  drop_na()

# export final dataset
write.csv(covid.clean, "covid.clean.csv")
```

# Data Management
```{r, warning=FALSE}
# check event var 
table(covid.clean$outcomec, covid.clean$event)

#check age variable
table(covid.clean$age_cat)

# remove orignal dataset
rm(list = "covid_org5")

covid.clean <- read.csv("covid.clean.csv")
```


#Plot survival time KM curve 
```{r}
ggplot(covid.clean[which(covid.clean$event==1),], aes(time, color=factor(country_binary), group=factor(country_binary))) + 
  geom_density(alpha=.1) #Histogram overlaid with kernel density curve for stage of diagnosis

ggplot(covid.clean[which(covid.clean$event==1),], aes(time, color=factor(sex), group=factor(sex))) + 
  geom_density(alpha=.1)

ggplot(covid.clean[which(covid.clean$event==1),], aes(time, color=factor(age_cat), group=factor(age_cat))) + 
  geom_density(alpha=.1)

```

#Get the KM values and plot KM curves for country & sex & age
```{r}
#KM curves
covid.clean.survcountry <- survfit(Surv(time, event) ~ country_binary, covid.clean) 
covid.sex.surv <- survfit(Surv(time, event) ~ sex, covid.clean)
covid.age.surv <- survfit(Surv(time, event) ~ age_cat, covid.clean)

#get KM values
summary(covid.clean.survcountry) 
summary(covid.sex.surv) 
summary(covid.age.surv)

#calculate KM plots with confidence intervals
ggsurvplot(covid.clean.survcountry, covid.clean = covid.clean, conf.int=TRUE, tables.theme = clean_theme())
ggsave("survcountr.png")

ggsurvplot(covid.sex.surv, data = covid.clean, conf.int=TRUE, tables.theme = clean_theme())
ggsave("survsex.png")

ggsurvplot(covid.age.surv, data = covid.clean, conf.int=T, tables.theme = clean_theme())
ggsave("survage.png")

```

#Determine the median survival time for each covariate 
```{r}
#Median survival
covid.clean.survcountry
covid.sex.surv
covid.age.surv
```

#log-rank test to determine any overall differences in covid-19 survival by country & sex & age categories
```{r}

survdiff(Surv(time, event) ~ country_binary,data=covid.clean)
survdiff(Surv(time, event) ~ sex,data=covid.clean)
survdiff(Surv(time, event) ~ age_cat,data=covid.clean)
```

```{r}
#pairwise comparison for age cateogry
pairwise_survdiff(Surv(time, event) ~ age_cat,data=covid.clean)

```

#Univariate Cox proportional hazards models
```{r}
#univariate Cox proportional hazards model binary country
treat.modcountry<-coxph(Surv(time, event)~country_binary, covid.clean, ties="efron") 

summary(treat.modcountry)

```

## Univaritate Cox PH Model for Sex
```{r}
cox.covid.sex <-coxph(Surv(time, event) ~ sex, data=covid.clean, ties="efron")

summary(cox.covid.sex)
```

## Univarite Cox PH Model for Age Categories
```{r}
cox.covid.age <-coxph(Surv(time, event) ~ age_cat, data=covid.clean, ties="efron")

summary(cox.covid.age)
```

#ADJUSTED Cox proportional hazards Model
```{r}

treat_adj.modcountry<-coxph(Surv(time, event)~country_binary + age_cat + sex, covid.clean)
summary(treat_adj.modcountry)

```


#Compare models
```{r}
#model comparison
library(lmtest)
lrtest(treat.modcountry, treat_adj.modcountry)

lrtest(cox.covid.sex, treat_adj.modcountry)

lrtest(cox.covid.age, treat_adj.modcountry)

```

#Check the PH assumption for your age and sex adjusted Cox model
```{r}
#Install and load packages
pacman::p_load(tidyverse, survival, survminer, survRM2, haven, gtools, magrittr, mlr)

#Check PH assumption using the Schoenfeld goodness-of-fit test 
#The global p-value tells whether the ph assumption is violated for any of the covariates. 
#You want this test to be non-significant. 

test.phcountry<-cox.zph(treat_adj.modcountry, terms=FALSE)
test.phcountry

test.phage <- cox.zph(cox.covid.age, terms=FALSE)
test.phage

```
Country violates PH assumption. 

#Run an univariate RMST analysis to calculate RMST and their difference
```{r}

#As the insurance variable has more than two levels, you need to use the rmst2_plus function that we created in class to estimate the RMST for all insurance groups, and their differences vs. private insurance
#Run the code below before further step
rmst2_plus <- function(data, time, status, arm, nlevel, ntime, ...){
  #nlevel: the number of levels of variable of interest minus 1
  for (i in 1:nlevel) {
    #level = 1, run the RMST model
    if (i==1) {
      print("########################################################")
      print(paste0("Below display the model results of arm =", i, " vs. arm = 0."))
      print(data %$%rmst2(time, status, arm, tau=ntime, ...))
    }
    #level > 1, change all levels to MA other than baseline level and the target level
    if (i>1) {
    print("#########################################################")
    print(paste0("Below display the model results of arm =", i, " vs. arm = 0."))
    print(
      data %>%
         mutate(arm=replace(arm, arm<i & arm>0, NA),
                arm=replace(arm, arm==i, 1)) %>%
         filter(!is.na(arm)) %$%
         rmst2(time, status, arm, tau=ntime, ...)
      )
     }
   }
}


#Choose the set of covariates 
covar <- c("sex2", "age_cat2")

modcountry <- rmst2(covid.clean$time, covid.clean$event, covid.clean$country_binary2, tau=40, covariates = covid.clean[, covar])
modcountry

modcountry.nocovar <- rmst2(covid.clean$time, covid.clean$event, covid.clean$country_binary2, tau=40)

#Plot the survival probability
plot(modcountry.nocovar, xlab="Days", ylab="Death Probability")
```



